R for Clinical Study Reports and Submission

### 2021-08-28 tlf-population.Rmd
tlf-population: fix typo

# Table for Study Patients {#population}

@@

In Chapter \@ref(dispositon), there are few repeating steps:

@@

```{r}
fmt_num <- function(x, digits, width = digits + 4){
  formatC(x,
          digits = digits, 
          format = "f", 
          width = width
    )
}
```

@@

For example, we can compare the two versions of code to format percentage 
used in \@ref(dispositon).

@@

```{r, eval = FALSE}
fmt_pct(n / n() * 100)
```

@@

  left_join(
    count(data, grp, var),
    count(data, grp, name = "tot"), 
    by = "grp",
  ) %>%
    mutate(
      pct = fmt_num(100 * n / tot, digits = 1),
      n   = fmt_num(n, digits = 0), 
      npct = paste0(n , " (", pct, ")")
    ) %>%
    pivot_wider(
      id_cols = var,
      names_from = grp,
      values_from = c(n, pct, npct),
      values_fill = list(n = "0", pct = fmt_pct(0))
    ) %>% 
    mutate(var_label = var_label)
}
```

@@

```{r}
count_by(adsl, "TRT01PN", "EFFFL")
```

### Analysis Code

@@

```{r}
# Derive a randomization flag 
adsl <- adsl %>% mutate(RANDFL = "Y") 
pop <- count_by(adsl, "TRT01PN", "RANDFL", var_label = "Participants in Population") %>% 
          select(var_label,starts_with("n_"))
```

```{r}
pop1 <- bind_rows(
  count_by(adsl, "TRT01PN", "ITTFL", 
           var_label = "Participants included in ITT population"), 
  count_by(adsl, "TRT01PN", "EFFFL", 
           var_label = "Participants included in efficacy population"),
  count_by(adsl, "TRT01PN", "SAFFL", 
           var_label = "Participants included in safety population")
) %>% 
  filter(var == "Y") %>% 
  select(var_label, starts_with("npct_"))
```

```{r}
names(pop) <- gsub("n_", "npct_", names(pop))
tbl_pop <- bind_rows(pop, pop1)
tbl_pop
```

```{r}
rel_width <- c(2, rep(1, 3))
colheader <- " | \\line Placebo | Xanomeline \\line Low Dose| Xanomeline \\line High Dose"
tbl_pop %>%

  # Table title
  rtf_title("Participants Accounting in Analysis Population", 
            "(All Participants Randomized)") %>%

  # First row of column header
  rtf_colheader(colheader,
    col_rel_width = rel_width
  ) %>%

  # Second row of column header
  rtf_colheader(" | n (%) | n (%) | n (%)",
    border_top = "",
    col_rel_width = rel_width) %>%

  # Table body
  rtf_body(
    col_rel_width = rel_width,
    text_justification = c("l", rep("c", 3))
  ) %>%

  # Encoding RTF syntax
  rtf_encode() %>%

  # Save to a file
  write_rtf("tlf/tbl_pop.rtf")
```

>>>>

# Table for Analysis Population {#population}

@@

In Chapter \@ref(disposition), there are few repeating steps to:

@@

```{r}
fmt_num <- function(x, digits, width = digits + 4) {
  formatC(x,
    digits = digits,
    format = "f",
    width = width
  )
}
```

@@

For example, we can compare the two versions of code to format percentage 
used in \@ref(disposition) and `fmt_num`.

@@

```{r, eval = FALSE}
fmt_num(n / n() * 100, digits = 1)
```

@@

  left_join(
    count(data, grp, var),
    count(data, grp, name = "tot"),
    by = "grp",
  ) %>%
    mutate(
      pct = fmt_num(100 * n / tot, digits = 1),
      n = fmt_num(n, digits = 0),
      npct = paste0(n, " (", pct, ")")
    ) %>%
    pivot_wider(
      id_cols = var,
      names_from = grp,
      values_from = c(n, pct, npct),
      values_fill = list(n = "0", pct = fmt_pct(0))
    ) %>%
    mutate(var_label = var_label)
}
```

@@

```{r}
count_by(adsl, "TRT01PN", "EFFFL") %>% 
  select(- ends_with(c("_54", "_81")))
```

## Analysis Code

@@

```{r}
# Derive a randomization flag
adsl <- adsl %>% mutate(RANDFL = "Y")

pop <- count_by(adsl, "TRT01PN", "RANDFL", 
                var_label = "Participants in Population") %>%
  select(var_label, starts_with("n_"))
```

```{r}
pop1 <- bind_rows(
  count_by(adsl, "TRT01PN", "ITTFL",
    var_label = "Participants included in ITT population"
  ),
  count_by(adsl, "TRT01PN", "EFFFL",
    var_label = "Participants included in efficacy population"
  ),
  count_by(adsl, "TRT01PN", "SAFFL",
    var_label = "Participants included in safety population"
  )
) %>%
  filter(var == "Y") %>%
  select(var_label, starts_with("npct_"))
```

Now we combined individual rows into the whole table for reporting purpose. 
`tbl_pop` is used as input for `r2rtf` to create final report.

```{r}
names(pop) <- gsub("n_", "npct_", names(pop))
tbl_pop <- bind_rows(pop, pop1)

tbl_pop %>% select(var_label, npct_0)
```

We start to define the format of the output.

```{r}
rel_width <- c(2, rep(1, 3))
colheader <- " | Placebo | Xanomeline line Low Dose| Xanomeline line High Dose"
tbl_pop %>%
  # Table title
  rtf_title(
    "Participants Accounting in Analysis Population",
    "(All Participants Randomized)"
  ) %>%
  # First row of column header
  rtf_colheader(colheader,
    col_rel_width = rel_width
  ) %>%
  # Second row of column header
  rtf_colheader(" | n (%) | n (%) | n (%)",
    border_top = "",
    col_rel_width = rel_width
  ) %>%
  # Table body
  rtf_body(
    col_rel_width = rel_width,
    text_justification = c("l", rep("c", 3))
  ) %>%
  # Encoding RTF syntax
  rtf_encode() %>%
  # Save to a file
  write_rtf("tlf/tbl_pop.rtf")
```
